---
title: "Calculate population level metrics for axis analysis"
author: "Kevin Healy"
date: "19 June 2017"
output:
  pdf_document: default
  html_document: default
---

This document outlines the analysis of how timing of mortailty is linked with 
parental care in the "Alternative measures of the shape of mortality reveals 
hidden type IV survivorship curves and associations with parental care at 
macroecological scales"

#Packages and Data
First we set up the libraries and data for the analysis. Packages such as 
mulTree will require using devtools to download from Github (See here https://github.com/TGuillerme/mulTree) 

```{r install_packages}
library(devtools)
library(popbio)
library(popdemo)
library(ape)
library(caper)
library(phytools)
library(MCMCglmm)
library(mulTree)
library(Rage)
```

The data is from the COMADRE database version 2.0.1. We will also upload some
useful functions fromt he COMADRE file.

```{r data}

load("COMADRE_v.X.X.X.RData")

source("COMADRE_functions.R")
```

The phylogeny is from the Healy et al 2019 paper (See here https://www.nature.com/articles/s41559-019-0938-7). We can plot it as well.

```{r phylogeny}

com_tree <- read.tree("COMADRE_phylo_june_2017_100.tre")
plot(com_tree[[1]], cex = 0.2, type = "fan")

```

Next we will  load some trait data associated with parental care.

```{r trait data}

trophic_data <- read.csv("Trophic_data_June16_2017.csv", 
                         sep = ",", header = T)

```


WE now need to subset the comadre database to the species we want. This follows
Healy et al 2019 (See here https://www.nature.com/articles/s41559-019-0938-7)
using pooled and mean metrices that are unmanipulated.

```{r subsetting}

###Subset entries form comadre with the below restictions for pooled and mean matrices.
pooled_Metadata <- (subset(comadre$metadata, 
                           MatrixComposite == "Pooled"
                           & MatrixDimension >= 2 
                           & StudyDuration >= 2 
                           & MatrixSplit == "Divided" 
                           & MatrixFec == "Yes" 
                           & MatrixTreatment == "Unmanipulated"
                           & AnnualPeriodicity == "1"
                           & SurvivalIssue<=1
))

mean_Metadata <- (subset(comadre$metadata,
                         MatrixComposite == "Mean" 
                         & MatrixDimension >= 2 
                         & StudyDuration >= 2 
                         & MatrixSplit == "Divided" 
                         & MatrixFec == "Yes" 
                         & MatrixTreatment == "Unmanipulated"
                         & AnnualPeriodicity == "1"
                         & SurvivalIssue<=1
))


###stick them together
combined_data <- rbind(pooled_Metadata, mean_Metadata)

###pull out the matching rows
keep_first <- as.numeric(rownames(combined_data))

##use these rows to pull out the matrices
combMat <- comadre$mat[keep_first]

#and associated matrix Class data
combmatrixClass <- comadre$matrixClass[keep_first]

##and set up a vector of the species
species_list <- data.frame(species =(combined_data$SpeciesAccepted), 
                           class = (combined_data$Class), 
                           phyla = (combined_data$Phylum))

#pull out the unique species
species_list_u <- unique(species_list$species)
```

Now check all matrices for egordicity, primativity and irriducablity using the
popdemo package. This will give us a list of clean species to use.

```{r check ergodicity, message=FALSE, warning=FALSE}

is_ergodic <- vector()
is_primitive <- vector()
is_irreducible <- vector()

all_true <- vector()

for(i in 1:length(keep_first)){
  tryCatch({
    is_ergodic[i] <- isErgodic(combMat[[i]]$matA)
    is_primitive[i] <- isPrimitive(combMat[[i]]$matA)
    is_irreducible[i] <- isIrreducible(combMat[[i]]$matA)

      }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  all_true[i] <- all(c(is_ergodic[i],is_primitive[i],is_irreducible[i]) == TRUE)
} 

test.frame <- data.frame(is_ergodic,is_primitive,is_irreducible)
keep <- which(all_true == TRUE)

discard_species <- which(all_true == FALSE)
discard_species_names  <- unique(combined_data[discard_species,]$SpeciesAccepted)

clean_species <- combined_data[keep,]$SpeciesAccepted

```


Now to calulate each of the metrics. First lets get some info on all the matrices we are using.

```{r species, population and taxa}

##loop through keep whaich are all the matces that passed the above tests.
##species names
species_ind_full <- vector()
##taxa name
taxa_name <- vector()
##population_name
pop_mat_name <- vector()
###name of the first stage in the matrix
first_stage <- vector() 
##matrix dimesnion
matrix_size <- vector()

  
for(i in 1:length(keep)){
  
##species names
species_ind_full[i] <-  combined_data[keep[i],]$SpeciesAccepted

##taxa name
taxa_name[i] <- as.vector(combined_data[keep[i],]$Class)

##population_name
pop_mat_name[i] <- combined_data[keep[i],]$MatrixPopulation

###name of the first stage in the matrix
first_stage[i] <- as.vector(combmatrixClass[keep[i]][[1]][1,2])
    
##matrix dimesnion
matrix_size[i] <- dim(combMat[[keep[i]]]$matA)[1]

}

first_stage <- unlist(first_stage)
```

Now lets calulate when only 5% and 1% are alive as a cutoff criteria, QSD and age at forst reporduction

```{r age at first reporduction}


##Age at first reproduction
life_time_La <- vector()

for(i in 1:length(keep)){
  tryCatch({
 
    ##### age at first reproduction
    life_time_La[i] <-	lifeTimeRepEvents(matU =   combMat[[keep[i]]]$matU, 
                                          matF =  combMat[[keep[i]]]$matF,
                                          startLife = 1)$La
    
    },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
```
    
Now lets calulate the age when 95% an 99% of the mature population is alive
```{r %alive}

##age (year) when 99% of chohort are dead
surv_99 <- vector()
##age (year) when 95% of chohort are dead
surv_95 <- vector()
#a resonable minimum maximum which is the hist of matrix size, 99% survival or QSD
min_max <- vector() 

for(i in 1:length(keep)){
  tryCatch({

    
    ##### age when 95% are dead
    surv_95[i] <- exceptionalLife(combMat[[keep[i]]]$matU, startLife = 1)[1]
    
    ##### age when 99% are dead
    
    surv_99[i] <- which((makeLifeTable(
                    matU = combMat[[keep[i]]]$matU, 
                    matF =  combMat[[keep[i]]]$matF, 
                    matC =  combMat[[keep[i]]]$matC, 
              startLife = 1, nSteps = 1000)$lx*1000) < 0.01)[1]
    

  # min_max[i] <- max(surv_99[i],matrix_size[i])
  min_max[i] <- surv_99[i] - 1
    
    },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

```



No let calulate some lifetables
```{r lifetables}

###life table
lxmx_curve <- list()

for(i in 1:length(keep)){
  tryCatch({

    ####calulculate the lxmx curve
    #I use max lifespan as the cut off point
    lxmx_curve[[i]] <-  makeLifeTable(matU =  combMat[[keep[i]]]$matU, 
                                 matF =  combMat[[keep[i]]]$matF, 
                                 matC =  combMat[[keep[i]]]$matC, 
                                 startLife = 1, nSteps = min_max[i])
  
  },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}   
```


k_entrpy and AUC
```{r mean_life_expect}
k_ent_rage <- vector()
AUC_rage <- vector()

k_ent_norm <- vector()
AUC_norm <- vector()

for(i in 1:length(keep)){
  tryCatch({	
    ##### mean life expectancy
    k_ent_rage[i] <-	entropy_k(lxmx_curve[[i]]$lx, trapeze = FALSE)
    
    AUC_rage[i] <-	shape_surv(lxmx_curve[[i]]$lx)
    
    k_ent_norm[i] <- kentropy_spline(lx = lxmx_curve[[i]]$lx, 
                                 x = lxmx_curve[[i]]$x,
                                 all.knots = TRUE,
                                 bins = 10)[1]
    
   AUC_norm[i] <- kentropy_spline(lx = lxmx_curve[[i]]$lx, 
                                 x = lxmx_curve[[i]]$x,
                                 all.knots = TRUE,
                                 bins = 10)[3]

     
        },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}   

```


Mean life expectancy
```{r mean_life_expect}
mean_life_expect <- vector()
for(i in 1:length(keep)){
  tryCatch({	
    ##### mean life expectancy
    mean_life_expect[i] <-	meanLifeExpectancy(matU = combMat[[keep[i]]]$matU, 
                                              startLife= 1)
     
        },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}   

```


Life expectancy conditional on reaching sexual maturity
```{r M_rep_lif_exp}
M_rep_lif_exp <- vector()
for(i in 1:length(keep)){
  tryCatch({	
 
  ##life expectancy conditional on reaching reporduction
    M_rep_lif_exp[i] <- lifeTimeRepEvents(matU =  combMat[[keep[i]]]$matU, 
                                          matF =  combMat[[keep[i]]]$matF,
                                          startLife = 1)$meanRepLifeExpectancy
     
        },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}   

```



standard diviation of survival
```{r Gini}


sbins <- 100

surv_sd <- vector()
surv_mean <- vector()

surv_sd_early <- vector()
surv_mean_early <- vector()

surv_sd_mature <- vector()
surv_mean_mature <- vector()
surv_early_sum <- vector()

prob_d_sd <- vector()
prob_d_sd_early <- vector()
prob_d_sd_mature <- vector()

fx_curve <- list()

prob_death <- list()

La_sbins <- vector()

spline_curve <- list()
##age at which 10% of the population are alive


##age at which 90% of the population are alive


##calulate f(x) age at death
for(i in 1:length(keep)){
  tryCatch({	
  
    
  spline_curve[[i]] <-  unlist(lx_spline(lxmx_curve[[i]]$lx, lxmx_curve[[i]]$x, bins = sbins)[1])
    
      La_sbins[i] <- round((life_time_La[i]/length(lxmx_curve[[i]]$lx))*sbins)

  
  
      for(z in 1:c(length(spline_curve[[i]])-1)){
    
    if(z == 1){fx_curve[i][1] <- 1 - c(spline_curve[[i]][z+1])}

    else{ fx_curve[[i]][z] <- c(spline_curve[[i]][z]) - spline_curve[[i]][z+1]
    
  }
      }
    
    
       for(z in 1:c(length(spline_curve[[i]])-1)){
    
    if(z == 1){prob_death[i][1] <- 1 - c(spline_curve[[i]][z+1])}

    else{ prob_death[[i]][z] <- 1 - c(spline_curve[[i]][z+1])/spline_curve[[i]][z]
    
  }
    }
    
    
      
       surv_sd[i] <- sd(fx_curve[[i]], na.rm = TRUE)
       prob_d_sd[i] <- sd(prob_death[[i]], na.rm = TRUE)

       surv_sd_early[i] <- sd(fx_curve[[i]][1:c(La_sbins[i]+1)], na.rm = TRUE)
       surv_sd_mature[i] <- sd(fx_curve[[i]][c(La_sbins[i]):c(length(fx_curve[[i]]))], na.rm = TRUE)
       
       prob_d_sd_early[i] <- sd(fx_curve[[i]][1:c(La_sbins[i]+1)], na.rm = TRUE)
       prob_d_sd_mature[i] <- sd(fx_curve[[i]][c(La_sbins[i]):c(length(fx_curve[[i]]))], na.rm = TRUE)
       
       
       
       surv_early_sum[i] <- 1 - lxmx_curve[[i]]$lx[La_sbins[i]+1]
       
       surv_sd_early[i] <- sd(fx_curve[[i]][1:c(La_sbins[i]+1)], na.rm = TRUE)

       
       surv_mean[i] <- mean(fx_curve[[i]], na.rm = TRUE)
       surv_mean_early[i] <- mean(fx_curve[[i]][1:c(La_sbins[i])], na.rm = TRUE)
       surv_mean_mature[i] <- mean(fx_curve[[i]][c(La_sbins[i]):c(length(fx_curve[[i]]))], na.rm = TRUE)
   
    
        },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}   

```




And put them all into a table
```{r table}
###some errors here could be cleaned up

mean_max_life <- mean_life_expect/surv_95

ind_vital <- data.frame(species_ind_full,
                        taxa_name,
                        pop_mat_name,
                        mean_life_expect,
                        mean_max_life,
                        life_time_La,
                        first_stage,
                        k_ent_rage,
                        AUC_rage,
                        k_ent_norm,
                        AUC_norm,
                        matrix_size
                        )

```

Now just remove the sub_species part of the names
```{r calulation of metrics at population level}

ind_vital[,"species_ind_full"]  <- gsub(" subsp.*","", ind_vital[,"species_ind_full"])
ind_vital[,"species_ind_full"]  <- gsub(" ","_", ind_vital[,"species_ind_full"])

```

Next we add in the external infor such as habitat etc.

```{r  trophic data}
##trophic data
mass_g <- vector()
repo_size_g <- vector()
par_care <- vector()
viv <- vector()

for(k in 1:(length(ind_vital$species_ind_full))){
  tryCatch({	

      mass_g[k] <-   (na.omit(trophic_data[trophic_data$species  == as.vector(ind_vital[k,"species_ind_full"]),"mass_g"]))[1]
      
            repo_size_g[k] <-   (na.omit(trophic_data[trophic_data$species  == as.vector(ind_vital[k,"species_ind_full"]),"repo_size_g"]))[1]
            
           
           par_care[k] <-   (na.omit(trophic_data[trophic_data$species  == as.vector(ind_vital[k,"species_ind_full"]),"parental_care"]))[1]


           viv[k] <-   (na.omit(trophic_data[trophic_data$species  == as.vector(ind_vital[k,"species_ind_full"]),"repo_type"]))[1]



      
      },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}


```


Stick them all together
```{r  trophic and pop data}

pop_vital <- data.frame(ind_vital,                       
                        mass_g,
                        repo_size_g,
                        par_care,
                        viv
                         )

```





Let also add some human demography from other sources external to COMADRE
```{r human data}
#load("PPMs Eusrostat 100x100.Rdata")
load("keyfitz_Agam_100x100.Rdata")
```


lets just calulate the whole lot together
```{r human metric}

#Spline bins

sbins <- 10

##I need to come up with a better breack down for these
#lets calculate for Ireland
Irish_rows_1960 <- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Ireland" & keyfitz$countryyear$year > 1960,]))
Irish_rows_1920 <- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Ireland" & keyfitz$countryyear$year < 1960,]))

#lets calculate for Sweden as Finland didnt work
Sweden_rows_a60 <- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Sweden" & keyfitz$countryyear$year > 1960,]))
Sweden_rows_1850_19 <- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Sweden" & keyfitz$countryyear$year > 1850 & keyfitz$countryyear$year < 1901,]))

#lets calculate for Japan
Japan_rows_b60 <- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Japan" & keyfitz$countryyear$year < 1960,]))
Japan_rows_a60 <- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Japan" & keyfitz$countryyear$year > 1960,]))


Irish_rows<- as.numeric(rownames(keyfitz$countryyear[keyfitz$countryyear$country == "Ireland",]))

##average the pops
Irish_mat_1960 <- keyfitz$Agam100[Irish_rows_1960]
Irish_mat_1920 <- keyfitz$Agam100[Irish_rows_1920]
Sweden_rows_1850_19 <- keyfitz$Agam100[Sweden_rows_1850_19]
Sweden_rows_a60 <- keyfitz$Agam100[Sweden_rows_a60]
japan_mat_b60 <- keyfitz$Agam100[Japan_rows_b60]
japan_mat_a60 <- keyfitz$Agam100[Japan_rows_a60]

Irish_1960 <- meanMatrix(Irish_mat_1960)
Irish_1920 <- meanMatrix(Irish_mat_1920)
Sweden_1850_19 <- meanMatrix(Sweden_rows_1850_19)
Sweden_a60 <- meanMatrix(Sweden_rows_a60)
japan_b60 <- meanMatrix(japan_mat_b60)
japan_a60 <- meanMatrix(japan_mat_a60)


##you need to put this in the right order
human_pop <- c("Homo_Irish1920" , "Homo_Irish1960", "Sweden_1850_19","Sweden_a60", "japan_b60", "japan_a60")
human_mats <- list(Irish_1920, Irish_1960, Sweden_1850_19, Sweden_a60, japan_b60, japan_a60)

##pop names and year
pop_country_year <- vector()
##mean life expectancy
h_mean_life_expect <- vector()
##age (year) when 99% of chohort are dead
h_surv_99 <- vector()
##age (year) when 95% of chohort are dead
h_surv_95 <- vector()
##life expectancy contingent on entering reproduction
h_M_rep_lif_exp <- vector()
##Age at first reproduction
h_life_time_La <- vector()
#generation time
h_gen_time <- vector()
##mean reporductive rate raw
h_mean_repo_rate <- vector()
##mean reporductive rate
h_mean_repo_rate_stable_state <- vector()
##time to quase stable distribution (how long does it take for the distrbution to stabalise on the Stable state distibution)
h_QSD <- vector()
#matrix dimension size
h_matrix_size <- vector()
h_gini <- vector()
h_mxlx_mean <- vector()
h_mxlxsd <- vector()
h_mxlxsd_prop <- vector()
h_mxlx_prop <- list()
h_mx_mean <- vector()
h_mxsd <- vector()
h_surv_sd <- vector()
h_surv_sd_early <- vector()
h_surv_sd_mature <- vector()
h_surv_mean <- vector()
h_surv_mean_early <- vector()
h_surv_mean_mature <- vector()

h_surv_early_sum <- vector()

fx_h_curve <- list()

h_lxmx_curve_list <- list()
h_spline_curve <- list()


h_k_ent_rage <- vector()
h_AUC_rage <- vector()

h_k_ent_norm <- vector()
h_AUC_norm <- vector()


##convert the age of maturity to the right point on the spline sampled curve
La_h_sbins <- vector()



for(i in 1:length(human_mats)){
  tryCatch({
    
###we first need to decompose the A matri into U F and C
    A_hum <-  human_mats[[i]]
    U_hum <- A_hum
    U_hum[1,] <- 0
    F_hum <- matrix(0,dim(A_hum)[1],dim(A_hum)[1])
    F_hum[1,] <- A_hum[1,]
    C_hum <- matrix(0,dim(A_hum)[1],dim(A_hum)[1])
    
    
    ##country and year
    pop_country_year[i] <- human_pop[i]

    ##matrix dimesnion
    h_matrix_size[i] <- dim(A_hum)[1]
    
    ### quasi-convergence for lx and mx 
    h_QSD[i] <- qsdConverge(U_hum, conv = 0.05, startLife = 1, 
                           nSteps = 10000)
     
    ##### mean life expectancy
    h_mean_life_expect[i] <-	meanLifeExpectancy(matU = U_hum, 
                                              startLife= 1)
    ##### age when 95% are dead
    h_surv_95[i] <- exceptionalLife(U_hum, startLife=1)[1]
    ##### age when 99% are dead
    h_surv_99[i] <- which(log10(makeLifeTable(
                    matU =  U_hum, 
                    matF =  F_hum, 
                    matC =  C_hum, 
              startLife = 1, nSteps = 1000)$lx*1000) < 0)[2]
    
    ##### time to first reproduction
    h_life_time_La[i] <-	lifeTimeRepEvents(matU = U_hum, 
                                          matF = F_hum,
                                          startLife = 1)$La
    
    ##life expectancy conditional on reaching reporduction
    h_M_rep_lif_exp[i] <- lifeTimeRepEvents(matU =  U_hum, 
                                          matF =  F_hum,
                                          startLife = 1)$meanRepLifeExpectancy
      

    ##### generation time from popbio package
    h_gen_time[i]  <- generation.time(A =  A_hum)
    
    ### mean reporduction rate
    h_mean_repo_rate_stable_state[i] <- meanRepo(matA = A_hum, 
                                                 matF = F_hum)
    
    h_mean_repo_rate[i] <- mean(F_hum[1,])
    
  
    
    ####calulculate the lxmx curve
    #I use max lifespan as the cut off point
    h_lxmx_curve <-  makeLifeTable(matU =  U_hum, 
                                 matF =  F_hum, 
                                 matC =  C_hum, 
                                 startLife = 1,
                                 nSteps = h_surv_99[i])
    
  
    
    La_h_sbins[i] <- round((h_life_time_La[i]/h_surv_99[i])*sbins)
    
    h_lxmx_curve_mature <-  makeLifeTable(matU =  U_hum, 
                                 matF =  F_hum, 
                                 matC =  C_hum, 
                                 startLife = h_life_time_La[i],
                                 nSteps = h_surv_99[i])
    

    h_gini[i] <- Gini(h_lxmx_curve$lx*h_lxmx_curve$mx)

  
  
    h_k_ent_rage[i] <-	entropy_k(h_lxmx_curve$lx, trapeze = FALSE)
    
    h_AUC_rage[i] <-	shape_surv(h_lxmx_curve$lx)

    h_surv_sd[i] <- sd(h_lxmx_curve$lx)
    
    h_k_ent_norm[i] <-  kentropy_spline(lx = h_lxmx_curve$lx, 
                                 x = h_lxmx_curve$x,
                                 all.knots = TRUE,
                                 bins = 10)[1]

    h_AUC_norm[i] <-  kentropy_spline(lx = h_lxmx_curve$lx, 
                                 x = h_lxmx_curve$x,
                                 all.knots = TRUE,
                                 bins = 10)[3]
      

    
    },error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}   


###some errors here could be cleaned up
h_ind_vital <- data.frame(                        
                        pop_country_year,
                        h_mean_life_expect,
                        h_life_time_La,
                        h_mean_repo_rate_stable_state,
                        h_mean_repo_rate,
                        h_gen_time,
                        h_gini,
                        h_M_rep_lif_exp,
                        h_matrix_size,
                        h_surv_95,
                        h_surv_99,
                        h_surv_sd,
                        h_k_ent_rage,
                        h_AUC_rage
                       )
    
```


```{r animal huma combine}


human_dem <- data.frame(species_ind_full = rep("Homo_sapiens", length(human_pop)), 
                        taxa_name =  rep("Mammalia" ,length(pop_country_year)),
                        pop_mat_name = pop_country_year,
                        mean_life_expect = h_mean_life_expect,
                        mean_max_life = c(h_mean_life_expect/h_surv_95),
                        life_time_La = h_life_time_La,
                        first_stage = rep("baby", length(human_pop)),
                        #surv_sd = h_surv_sd,
                        k_ent_rage = h_k_ent_rage,
                        AUC_rage = h_AUC_rage,
                        k_ent_norm = h_k_ent_norm,
                        AUC_norm = h_AUC_norm,
                        matrix_size = h_matrix_size,
                        mass_g =  rep(trophic_data[trophic_data$species  == "Homo_sapiens","mass_g"],length(pop_country_year)),
                        repo_size_g = rep(trophic_data[trophic_data$species  == "Homo_sapiens","repo_size_g"],length(pop_country_year)),
                        par_care = rep(trophic_data[trophic_data$species  == "Homo_sapiens","parental_care"],length(pop_country_year)),
                        viv = rep(trophic_data[trophic_data$species  == "Homo_sapiens","repo_type"],length(pop_country_year))
                        )

full_m_data <- rbind(pop_vital,
                     human_dem)

#add mass specific 

full_m_data <- cbind(full_m_data, 
                     repo_mas_sp = full_m_data$repo_size_g/full_m_data$mass_g)

```


Now we can clean the data before expoting it


Remove biologically crazy species

```{r remove biologically crazy species}

pop_data <- full_m_data

####remonve these problem species as they clearly have a problem
##Spratt have lifesspans >100 which clearly way off any sensible biology
pop_data_sr <- pop_data[pop_data$species != "Sprattus_sprattus",]

##This stupid monkey thinks it can live for >105
pop_data_sr2 <- pop_data_sr[pop_data_sr$species != "Brachyteles_hypoxanthus",]

##Scolytus_ventralis is coded in as annual but its clearyl a seasonal species
pop_data_sr3 <- pop_data_sr2[pop_data_sr2$species != "Scolytus_ventralis",]

####remove Enhydra_lutris its a simulation based study
##This gives an Na at hte moment, check at the end to see if it still does, in which case remove this line
pop_data_sr4 <- pop_data_sr3[pop_data_sr3$species != "Enhydra_lutris",]

pop_data_sr5 <- pop_data_sr4[pop_data_sr4$species != "Somateria_mollissima",]


####remove Chrysemys_picta as the fecundity is higher than the max reporduction in the paper
pop_data_sr6 <- pop_data_sr5[pop_data_sr5$species != "Chrysemys_picta",]
pop_data_sr7 <- pop_data_sr6[pop_data_sr6$species != "Crocodylus_johnsoni",]
#pop_data_sr8 <- pop_data_sr7[pop_data_sr7$species != "Amphimedon_compressa",]



###remove Mya_arenaria and Haliotis_laevigata as they have net repoductie rates of > 100
#full_nl_data <- full_nl_data[full_nl_data$species != "Amphimedon_compressa",]
#full_nl_data <- full_nl_data[full_nl_data$species != "Haliotis_laevigata",]

##remove the population colunm
drops <- c("pop_mat_name")
pop_data_nopop <- pop_data_sr7[,!(names(pop_data_sr7) %in% drops)]

#rename the species_ind_full col name to "species"
colnames(pop_data_nopop)[1] <- "species"
```


Now remove any infs of Na's in the pop metrics
```{r remove rows with inf or NAs}

##remove infinaties
pop_data_infr  <- do.call(data.frame,lapply(pop_data_nopop, function(x) replace(x, is.infinite(x),NA)))

full_infna_data <- na.omit(pop_data_infr)
```

```{r prun the tree down the remaining species}

##add a colume to idenfy pgls match up
phylo_match_data <- data.frame(full_infna_data, animal = full_infna_data$species)

pgls_unique_matched <- data.frame(species = unique(phylo_match_data$species), 
                                  species_pgls = unique(phylo_match_data$animal))


axis_trees <- list()
for(i in 1:(length(com_tree))){
tree_ren <- com_tree[[i]]

#tree match
axis_comp<- comparative.data(phy = tree_ren, 
                             data = pgls_unique_matched, 
                             names.col = "species_pgls" , force.root = TRUE)
axis_trees[[i]] <- axis_comp$phy
}



#small fix on ultrametric problem
is_ultra <- vector()
for(i in 1:length(axis_trees)) { 
  is_ultra[i] <- is.ultrametric(axis_trees[[i]])
  }

ultra_row <- which(is_ultra ==TRUE)

axis_trees <- axis_trees[ultra_row]

class(axis_trees) <- "multiPhylo"



axis_comp$dropped$unmatched.rows
```

First we make a multree object so we can loop each of the models through the trees
```{r mulTree}
pop_multree <- as.mulTree(data = phylo_match_data, tree = axis_trees, taxa = "animal", rand.terms = ~animal + species)

```

Now lets set a prior
```{r MCMC prior}

prior<-list(R = list(V = 1/2, nu=0.002), 
            G = list(G1=list(V = 1/2,n = 1, alpha.mu=rep(0,1), alpha.V= diag(1)*10^3), 
                     G1=list(V = 1/2,n = 1, alpha.mu=rep(0,1), alpha.V= diag(1)*10^3)))

```

and general parameters
```{r MCMC parameters}

#parameters <- c(1100000, 500, 100000)
parameters <- c(110000, 50, 10000)

```



#AUC_rage

```{r AUC_rage mulTree run}
formula_AUC_rage <- AUC_rage ~ par_care

mulTree(mulTree.data = pop_multree,
        formula = formula_AUC_rage,
        priors = prior,
        parameters = parameters,
        output = "AUC_rage_run",
        ESS = 1000,
        chains = 2)
```


All the diagnostices look good, lets read that back in.
```{r AUC_rage readback run}
AUC_rage_models <- read.mulTree("AUC_rage_run")
summary(AUC_rage_models)
```

life_shape_models variance terms

```{r AUC_rage var}
AUC_rage_var <- read.mulTree("AUC_rage_run", extract = "VCV")

AUC_rage_phlyo <- list()
AUC_rage_spec <- list()
AUC_rage_unit <- list()

for(i in 1:length(names(AUC_rage_var))){

  AUC_rage_phlyo[[i]] <-  AUC_rage_var[[1]][,1] 
  AUC_rage_spec[[i]] <-  AUC_rage_var[[1]][,2] 
  AUC_rage_unit[[i]] <-  AUC_rage_var[[1]][,3] 
  }

AUC_rage_phlyo <- unlist(AUC_rage_phlyo)
AUC_rage_spec <- unlist(AUC_rage_spec)
AUC_rage_unit <- unlist(AUC_rage_unit)

AUC_rage_prop_phlyo <- AUC_rage_phlyo/(AUC_rage_phlyo + AUC_rage_spec + AUC_rage_unit)
AUC_rage_prop_spec  <- AUC_rage_spec/(AUC_rage_phlyo + AUC_rage_spec + AUC_rage_unit)
AUC_rage_prop_residual  <- AUC_rage_unit/(AUC_rage_phlyo + AUC_rage_spec + AUC_rage_unit)

hdr(AUC_rage_prop_phlyo)
hdr(AUC_rage_prop_spec)
hdr(AUC_rage_prop_residual)

```


```{r AUC_rage plot}

##Allometric scaling
#pdf("scaling_bar_plots.pdf")
AUC_rage_list <- list(AUC_rage_prop_residual,
                      AUC_rage_prop_phlyo,
                      AUC_rage_prop_spec,
                      AUC_rage_models$par_careyes,
                      AUC_rage_models$`(Intercept)`
                      )


MultiDisPlot(AUC_rage_list)
abline(v = 0,
       lty = 3)

#dev.off()


```




#k_ent_rage

```{r k_ent_rage mulTree run}
formula_k_ent_rage <- k_ent_rage ~ par_care

mulTree(mulTree.data = pop_multree,
        formula = formula_k_ent_rage,
        priors = prior,
        parameters = parameters,
        output = "k_ent_rage_run",
        ESS = 1000,
        chains = 2)
```


All the diagnostices look good, lets read that back in.
```{r k_ent_rage readback run}
k_ent_rage_models <- read.mulTree("k_ent_rage_run")
summary(k_ent_rage_models)
```

life_shape_models variance terms

```{r k_ent_rage var}
k_ent_rage_var <- read.mulTree("k_ent_rage_run", extract = "VCV")

k_ent_rage_phlyo <- list()
k_ent_rage_spec <- list()
k_ent_rage_unit <- list()

for(i in 1:length(names(k_ent_rage_var))){

  k_ent_rage_phlyo[[i]] <-  k_ent_rage_var[[1]][,1] 
  k_ent_rage_spec[[i]] <-  k_ent_rage_var[[1]][,2] 
  k_ent_rage_unit[[i]] <-  k_ent_rage_var[[1]][,3] 
  }

k_ent_rage_phlyo <- unlist(k_ent_rage_phlyo)
k_ent_rage_spec <- unlist(k_ent_rage_spec)
k_ent_rage_unit <- unlist(k_ent_rage_unit)

k_ent_rage_prop_phlyo <- k_ent_rage_phlyo/(k_ent_rage_phlyo + k_ent_rage_spec + k_ent_rage_unit)
k_ent_rage_prop_spec  <- k_ent_rage_spec/(k_ent_rage_phlyo + k_ent_rage_spec + k_ent_rage_unit)
k_ent_rage_prop_residual  <- k_ent_rage_unit/(k_ent_rage_phlyo + k_ent_rage_spec + k_ent_rage_unit)

hdr(k_ent_rage_prop_phlyo)
hdr(k_ent_rage_prop_spec)
hdr(k_ent_rage_prop_residual)

```


```{r k_ent_rage plot}

##Allometric scaling
#pdf("scaling_bar_plots.pdf")
k_ent_rage_list <- list(k_ent_rage_prop_residual,
                      k_ent_rage_prop_phlyo,
                      k_ent_rage_prop_spec,
                      k_ent_rage_models$par_careyes,
                      k_ent_rage_models$`(Intercept)`
                      )


MultiDisPlot(k_ent_rage_list)
abline(v = 0,
       lty = 3)

#dev.off()


```


#mean_max_life

```{r mean_max_life mulTree run}
formula_mean_max_life <- mean_max_life ~ par_care

mulTree(mulTree.data = pop_multree,
        formula = formula_mean_max_life,
        priors = prior,
        parameters = parameters,
        output = "mean_max_life_run",
        ESS = 1000,
        chains = 2)
```


All the diagnostices look good, lets read that back in.
```{r mean_max_life readback run}
mean_max_life_models <- read.mulTree("mean_max_life_run")
summary(mean_max_life_models)
```

life_shape_models variance terms

```{r mean_max_life var}
mean_max_life_var <- read.mulTree("mean_max_life_run", extract = "VCV")

mean_max_life_phlyo <- list()
mean_max_life_spec <- list()
mean_max_life_unit <- list()

for(i in 1:length(names(mean_max_life_var))){

  mean_max_life_phlyo[[i]] <-  mean_max_life_var[[1]][,1] 
  mean_max_life_spec[[i]] <-  mean_max_life_var[[1]][,2] 
  mean_max_life_unit[[i]] <-  mean_max_life_var[[1]][,3] 
  }

mean_max_life_phlyo <- unlist(mean_max_life_phlyo)
mean_max_life_spec <- unlist(mean_max_life_spec)
mean_max_life_unit <- unlist(mean_max_life_unit)

mean_max_life_phlyo_prop <- mean_max_life_phlyo/(mean_max_life_phlyo + mean_max_life_spec + mean_max_life_unit)
mean_max_life_spec_prop  <- mean_max_life_spec/(mean_max_life_phlyo + mean_max_life_spec + mean_max_life_unit)
mean_max_life_residual_prop  <- mean_max_life_unit/(mean_max_life_phlyo + mean_max_life_spec + mean_max_life_unit)

hdr(mean_max_life_phlyo_prop)
hdr(mean_max_life_spec_prop)
hdr(mean_max_life_residual_prop)

```



#AUC_norm

```{r AUC_norm mulTree run}
formula_AUC_norm <- AUC_norm ~ par_care

mulTree(mulTree.data = pop_multree,
        formula = formula_AUC_norm,
        priors = prior,
        parameters = parameters,
        output = "AUC_norm_run",
        ESS = 1000,
        chains = 2)
```


All the diagnostices look good, lets read that back in.
```{r AUC_norm readback run}
AUC_norm_models <- read.mulTree("AUC_norm_run")
summary(AUC_norm_models)
```

AUC_norm variance terms

```{r AUC_norm var}
AUC_norm_var <- read.mulTree("AUC_norm_run", extract = "VCV")

AUC_norm_phlyo <- list()
AUC_norm_spec <- list()
AUC_norm_unit <- list()

for(i in 1:length(names(AUC_norm_var))){

  AUC_norm_phlyo[[i]] <-  AUC_norm_var[[1]][,1] 
  AUC_norm_spec[[i]] <-  AUC_norm_var[[1]][,2] 
  AUC_norm_unit[[i]] <-  AUC_norm_var[[1]][,3] 
  }

AUC_norm_phlyo <- unlist(AUC_norm_phlyo)
AUC_norm_spec <- unlist(AUC_norm_spec)
AUC_norm_unit <- unlist(AUC_norm_unit)

AUC_norm_phlyo_prop <- AUC_norm_phlyo/(AUC_norm_phlyo + AUC_norm_spec + AUC_norm_unit)
AUC_norm_spec_prop  <- AUC_norm_spec/(AUC_norm_phlyo + AUC_norm_spec + AUC_norm_unit)
AUC_norm_residual_prop  <- AUC_norm_unit/(AUC_norm_phlyo + AUC_norm_spec + AUC_norm_unit)

hdr(AUC_norm_phlyo_prop)
hdr(AUC_norm_spec_prop)
hdr(AUC_norm_residual_prop)

```



#k_ent_norm

```{r k_ent_norm mulTree run}
formula_k_ent_norm <- k_ent_norm ~ par_care

mulTree(mulTree.data = pop_multree,
        formula = formula_k_ent_norm,
        priors = prior,
        parameters = parameters,
        output = "k_ent_norm_run",
        ESS = 1000,
        chains = 2)
```


All the diagnostices look good, lets read that back in.
```{r AUC_norm readback run}
k_ent_norm_models <- read.mulTree("k_ent_norm_run")
summary(k_ent_norm_models)
```

k_ent_norm variance terms

```{r AUC_norm var}
k_ent_norm_var <- read.mulTree("k_ent_norm_run", extract = "VCV")

k_ent_norm_phlyo <- list()
k_ent_norm_spec <- list()
k_ent_norm_unit <- list()

for(i in 1:length(names(k_ent_norm_var))){

  k_ent_norm_phlyo[[i]] <-  k_ent_norm_var[[1]][,1] 
  k_ent_norm_spec[[i]] <-  k_ent_norm_var[[1]][,2] 
  k_ent_norm_unit[[i]] <-  k_ent_norm_var[[1]][,3] 
  }

k_ent_norm_phlyo <- unlist(k_ent_norm_phlyo)
k_ent_norm_spec <- unlist(k_ent_norm_spec)
k_ent_norm_unit <- unlist(k_ent_norm_unit)

k_ent_norm_prop_phlyo <- k_ent_norm_phlyo/(k_ent_norm_phlyo + k_ent_norm_spec + k_ent_norm_unit)
k_ent_norm_prop_spec  <- k_ent_norm_spec/(k_ent_norm_phlyo + k_ent_norm_spec + k_ent_norm_unit)
k_ent_norm_prop_residual  <- k_ent_norm_unit/(k_ent_norm_phlyo + k_ent_norm_spec + k_ent_norm_unit)

hdr(k_ent_norm_prop_phlyo)
hdr(k_ent_norm_prop_spec)
hdr(k_ent_norm_prop_residual)

```




##PLots
```{r plots}

par(mfrow=c(1,2))

plot(pop_multree$data$AUC_rage ~ as.factor(pop_multree$data$par_care),
     col ="darkgrey",
     bty = "n",
     boxwex = 0.05,
     ylab = "AUC",
     xlab = "Parental care")

points(pop_multree$data$AUC_rage ~ jitter(as.integer(as.factor(pop_multree$data$par_care)), 
                                          amount = 0.1),
       pch = 16,
       col = rgb(44,137,160, max=255),
       cex = 0.7)


points(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "AUC_rage"] ~ rep(2, length(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "par_care"])) , 
       pch = 16,
       col = "red")

points(pop_multree$data[pop_multree$data$species == "Oncorhynchus_tshawytscha", "AUC_rage"] ~ rep(1, length(pop_multree$data[pop_multree$data$species == "Oncorhynchus_tshawytscha", "par_care"])) , 
       pch = 16,
       col = "green")



plot(pop_multree$data$k_ent_rage ~ as.factor(pop_multree$data$par_care),
     col ="darkgrey",
     bty = "n",
     boxwex = 0.05,
     ylab = "H-entropy",
     xlab = "Parental care")

points(pop_multree$data$k_ent_rage ~ jitter(as.integer(as.factor(pop_multree$data$par_care)), 
                                          amount = 0.1),
       pch = 16,
       col = rgb(44,137,160, max=255),
       cex = 0.7)

points(pop_multree$data[pop_multree$data$species == "Oncorhynchus_tshawytscha", "k_ent_rage"] ~ rep(1, length(pop_multree$data[pop_multree$data$species == "Oncorhynchus_tshawytscha", "par_care"])) , 
       pch = 16,
       col = "green")


points(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "k_ent_rage"] ~ rep(2, length(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "par_care"])) , 
       pch = 16,
       col = "red")

```






##PLots
```{r plots}

par(mfrow=c(1,2))

plot(pop_multree$data$AUC_norm ~ as.factor(pop_multree$data$par_care),
     col ="darkgrey",
     bty = "n",
     boxwex = 0.05,
     ylab = "AUC",
     xlab = "Parental care")

points(pop_multree$data$AUC_norm ~ jitter(as.integer(as.factor(pop_multree$data$par_care)), 
                                          amount = 0.1),
       pch = 16,
       col = rgb(44,137,160, max=255),
       cex = 0.7)


points(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "AUC_norm"] ~ rep(2, length(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "par_care"])) , 
       pch = 16,
       col = "red")

points(pop_multree$data[pop_multree$data$species == "Chelydra_serpentina", "AUC_norm"] ~ rep(1, length(pop_multree$data[pop_multree$data$species == "Chelydra_serpentina", "par_care"])) , 
       pch = 16,
       col = "green")



plot(pop_multree$data$k_ent_norm ~ as.factor(pop_multree$data$par_care),
     col ="darkgrey",
     bty = "n",
     boxwex = 0.05,
     ylab = "H-entropy",
     xlab = "Parental care")

points(pop_multree$data$k_ent_norm ~ jitter(as.integer(as.factor(pop_multree$data$par_care)), 
                                          amount = 0.1),
       pch = 16,
       col = rgb(44,137,160, max=255),
       cex = 0.7)

points(pop_multree$data[pop_multree$data$species == "Chelydra_serpentina", "k_ent_norm"] ~ rep(1, length(pop_multree$data[pop_multree$data$species == "Chelydra_serpentina", "par_care"])) , 
       pch = 16,
       col = "green")


points(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "k_ent_norm"] ~ rep(2, length(pop_multree$data[pop_multree$data$species == "Homo_sapiens", "par_care"])) , 
       pch = 16,
       col = "red")


```



